---
title: "lab_02"
author: "derek willis"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1** Create a new column that divides the expense amounts into the following categories - \$0-\$500, \$500-\$5000, \$5001-\$50000, \$50001\_\$100000, \$100001-\$500000 - giving each an appropriate text value, and count how many expenses are in each group. How would you describe the distribution of expenses in a sentence? Would you change the categories?

**A1.** The distributions decreased, as the expenses range increased in total. No, I would not change the categories because it's an easy way to comprehend the data.

```{r} 
maryland_expenses_with_state <- maryland_expenses %>%
  mutate(
    category = case_when(
      amount <500 ~ "0-$500",
      # I want to know the answers between 500 & 5000
      amount>= 500 & amount<5000 ~"500-5000",
      # I want to know the answers between 5001 & 50000
      amount>= 5001 & amount<50000 ~ "5001-50000",
      # I want to know the answers between 50001 & 100000
      amount>= 50001 & amount<100000 ~ "50001-1000000",
      # I want to know the answers between 100001 & 500000
      amount>= 100001 & amount<500000 ~ "100001-500000"
)
) %>% 
    group_by(category) %>%
  summarize(total = n()) %>%
  arrange(desc(total))
```

**Q2.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? 
**A2.** Atmospheric and Oceanic Science

```{r}

umd_courses %>% 
  filter(str_detect(title, "Climate")) %>% filter(seats>1) %>% 
  group_by(department) %>% 
  summarize(total = n()) %>%
  arrange(desc(total))










```

**Q3.** Of the departments listed in A2, which one is the most interesting to you, and why?
**A3.** Journalism. It's interesting because you wouldn't think climate plays a role.

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.**
As the election got closer, he spent more money on his campaign. The further away from the election, the less they spent.
```{r}
maryland_expenses <- maryland_expenses %>% mutate(expenditure_date=mdy(expenditure_date))
```

```{r}
maryland_expenses %>% 
  filter(committee_name== "Moore  Wes For Maryland") %>% 
  mutate(month = floor_date(expenditure_date, "month")) %>%
  group_by(month,committee_name) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```




**Q5.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.

What JOUR course is taught the most?

JOUR 175: Media Literacy is taught the most with 21 total classes.


**A5.** 

```{r}
umd_courses %>%
  filter(department=="Journalism") %>% 
  group_by(id) %>%
  summarise(total_classes = n()) %>%
  arrange(desc(total_classes))



```
