---
title: "lab_08"
author: "derek willis"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. We'll be making some charts, working with dates and retrieving Census data.
```{r}
library(tidyverse)
library(lubridate)
library(tidycensus)
```


## Load and modify data

**Task** Create a codeblock and load the following data from the data folder:

* Prince George's County 911 Overdose Calls

You will need to create columns for the date, week and month based on the existing `datetime` column.
```{r}
pgco_911_calls <- read_csv ("data/prince_georges_2022_overdoses.csv") %>% 
mutate(month=month(datetime)) %>% 
mutate(date=date(datetime)) %>% 
mutate(week=week(datetime)) 
```
group_by (month) %>% 
summarize(total_calls = n ()) %>% 
arrange(desc(total_calls))



## Questions 

**Q1.** Which month saw the greatest percentage of total calls? Create a dataframe that calculates the percentage of all calls that each month's calls represents. Make a bar chart from that dataframe. Your bar chart must have:

* A clear title that states the main idea/finding
* Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
* Readable bars - the values shouldn't be overlapping

Compare the results here to those from the pre_lab_08 bar chart - are there major differences in the months with the highest and lowest figures? Describe that below.

**A1.** December saw the greatest percentage of total calls with 10.1%. Yes, there are a few major differences from the pre_lab_08 bar chart to the bar chart created in Lab 08. The pre_lab_08 bar chart has August, May, March, and April as the top four months with the most total calls. While Lab 08 has December, July, November, and September as their top four months with the most calls. The four months with the lowest total calls in pre_lab_08 are January, September, December, and February. While the lowest total for Lab 08 are in the months of June, August, February, and January. 

```{r}
pgco_911_calls  %>% 
group_by(month) %>% 
summarize(total_calls = n()) %>% 
mutate(percentage= total_calls/1397*100) %>% 
arrange(desc(total_calls)) %>% 
ggplot()   +
  geom_bar(aes(x=reorder(month, total_calls), weight=total_calls)) +
  coord_flip()   +
  theme_minimal()
```

**Q2.** Let's visualize this data on a weekly basis using a line chart. As in Q1, generate a dataframe with the total number of calls for each week, and then create a line chart to show the distribution of calls over time. Your line chart must have:

* A clear title that states the main idea/finding
* Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
* Readable labels

Describe the pattern of results; you may want to look at the data to dig into particular outliers.

**A2.**The graph made its first impression in the beginning, middle, and end. Towards the end of the graph, it is pointing down. The end of the graph symbolizes the end of the year. The numbers are deflated because Dec. 31 landed on a Monday, which is the beginning of the week. It is difficult and unreasonable to confirm results with one day of data.


```{r}
pgco_911_calls  %>% 
  group_by(week) %>% 
summarize(total_calls = n()) %>% 
ggplot()  +
  geom_line(aes(x=week, y=total_calls)) +
labs(
  title = "Total Number of Overdose Calls in PG County",
  x="month",
  y="total_calls",
  caption ="source:Prince George's County EMS",
)
```

**Q3.**  A scatterplot is a type of chart that helps us see relationships between two variables. One variable goes on the x axis, the other on the y axis.  For each row/observation in our data, a scatterplot puts a circle (or a "point") where the two variables intersect on a grid. 

Statisticians use scatterplots to show graphically whether one variable is correlated -- related, in a statistical sense -- with another variable.  A classic example is the [relationship between ice cream sales and temperature](https://www.mathsisfun.com/data/scatter-xy-plots.html). The scatterplot below -- press play to load the image -- shows that relationship, that an increase in temperature is associated with an increase in ice cream sales. When it's 12C, sales are 200 dollars, and when it's hotter, 25C, sales are 600 dollars.

```{r}
knitr::include_graphics("https://www.mathsisfun.com/data/images/scatter-ice-cream1.svg")
```

We're going to use a scatterplot a little differently, to get a visual sense of two key variables: 

Our question is: does the median income in a zip code have any relationship to the number of overdose 911 calls in that zip code?

To answer this question, do the following:

1. Generate a dataframe with the number of 911 calls for each zip code.
2. Get data from the Census Bureau showing median household income for Maryland zip codes.
3. Join those two dataframes on their zip code columns, starting with the 911 calls dataframe.
4. Make a scatterplot showing the total calls and median income. I didn't show you how to do this, so look it up! Googling "ggplot scatterplot" is a good start.
5. Give it an appropriate title, source, and x and y axis titles.
6. Add a label for each point that shows the zip code using geom_text() - see some examples of its use at https://ggplot2.tidyverse.org/reference/geom_text.html#ref-examples. Try to make the names as easy to read as possible by avoiding overlap.
7. In the answer space below, describe what you see and answer the questions posed above. In a general sense, what do you think this means? Feel free to consider the actual raw values: how would you report out the main point(s) of this chart?

**A3.**I believe this chart represents the total number of calls over a month and where they’re coming from. I would report out that the chart will show where calls are coming from and how frequently they are coming in over the span of a month.  

```{r}
pgzip_911 <- pgco_911_calls  %>% 
  group_by(zipcode) %>% 
summarize(total_calls = n())
pgco <- get_acs(geography = "zcta",
              variables = c(medincome = "B19013_001"),
              state = "MD",
              year = 2019)
 pgcototal_911 <- pgzip_911 %>% left_join(pgco, by=c('zipcode'='GEOID'))

 pgcototal_911 %>% 
ggplot()   +
   geom_point(aes(x=estimate, y=total_calls)) +
   labs(
  title = "Total Number of Overdose Calls in PG County",
  x="month",
  y="total_calls",
  caption ="source:Prince George's County EMS",
  )
 geom_text(check_overlap = TRUE) +
 geom_label() 
 
 
```
