---
title: "lab_11"
author: "derek willis"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries for working with data, plus rvest.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
#| output: false
library(rvest)
library(tidyverse)
library(janitor)
library(lubridate)
```



Let's get to scraping.

## Questions

**Q1**. Scrape the listing of available Maryland state grants at https://grants.maryland.gov/Pages/StateGrants.aspx into a dataframe. You should have three columns, one of which is a date, so make sure the date column has a date datatype. Then, write code to count the number of grants opportunities offered by each organization listed in your dataframe, showing the organization with the most grant opportunities first. Which state agency has the most?

**A1**The Maryland Energy Administration was the state agency with the most grant opportunities with 14.One thing that I found interesting was that the Department of Commerce and Department of Health and Mental Hygiene only had one grant opportunity. The other thing that I noticed was that there were duplicate names of organizations, but it was written differently. For example: the first entry is written as “Maryland Energy Administration.” While the tenth entry is written as “MD Energy Administration.” 

```{r}
# scrape MD state grants as a dataframe
grants_url <- "https://grants.maryland.gov/Pages/StateGrants.aspx"
```

```{r}
# read in the html
results <- grants_url %>%
  read_html()

# display the html below
results
```
```{r}
# Read in all html from table, store all tables on page as nested list of dataframes.
results <- grants_url %>%
  read_html() %>%
  html_table()

# Just keep the first dataframe in our list

results <- results[[1]]

# show the dataframe

results
```


```{r}
# read in the html and extract all the tables
results <- grants_url %>%
  read_html() %>%
  html_table()

# show the dataframe

results
```
```{r}
# Read in all html from table, get the HTML table.
results <- grants_url %>%
  read_html() %>%
  html_table()

# Standardize column headers, remove last row

results <- results[[1]] %>%
  clean_names() %>%
  mutate(due_date = mdy(due_date))

# show the dataframe
results
```
```{r}
#Then, write code to count the number of grants opportunities offered by each organization listed in your dataframe
results %>%
  group_by(organization) %>% 
  summarize(count= n())  %>% 
  arrange(desc(count)) 
```




**Q2** Next, let's scrape the list of press releases from Maryland's Office of the Public Defender, https://www.opd.state.md.us/press-releases. This isn't a table, so you'll need to use `html_elements()` and your browser's inspector and do some clean up on the results. The result should be a dataframe with two columns that contain the date and title, and the date column should have a date datatype. The challenge here is figuring out how to isolate the releases.

When you finish scraping into a dataframe, write code to find the press releases that have the word "police" in the title. How many are there and when was the most recent one?

**A2**There are nine entries with the most recent being June 21, 2021.

```{r}
pressreleases_url <- "https://www.opd.state.md.us/press-releases"
```

```{r}
# read in the html
results <- pressreleases_url %>%
  read_html()

# display the html below
results
```
```{r}
results %>% html_elements('li')
```
```{r}
results %>% html_elements('p') %>% html_text()
```
```{r}
pressreleases <- results %>% html_elements('p') %>% html_text() %>% as_tibble()
```

```{r}
pressreleases <- results %>% html_elements('a.wixui-rich-text__text') %>% html_text() %>% as_tibble()
```

```{r}
pressreleases <- results %>% html_elements('a.wixui-rich-text__text') %>% html_text() %>% as_tibble() %>% separate(value, c('date', 'title'), sep=":")
```
```{r}
pressreleases <- results %>% html_elements('a.wixui-rich-text__text') %>% html_text() %>% as_tibble() %>% separate(value, c('date', 'title'), sep=":") %>% drop_na(title) %>% mutate(title = str_squish(title), date = mdy(date))

```
```{r}
pressreleases %>% 
    filter(str_detect(title, "Police"))
```








**Q3** Sen. Ben Cardin, D-Maryland, has posted hundreds of press releases at https://www.cardin.senate.gov/?post_type=press-releases. It would be great to have all of them in a dataframe that has the following columns: date, title and url.

To do this, you will need to scrape the page's html and save that to a variable, and _then_ extract the dates, titles and urls into _separate_ dataframes using html_elements(). And remember how we turn a list into a dataframe. The function `html_text()` pulls out the contents of a tag, but for urls we want the HTML attribute. Rvest gives you a way to extract the URL from a link; google to find out what it is.

At the end, you'll have three dataframes that you want to combine into a single dataframe. When we want to combine the rows of identical dataframes, we used `bind_rows()`. If you were combining columns instead of rows, there's a similar function. Find out what it is and use it to put all of the dataframes together into a single one.

When you're done, rename the columns so they make sense, then make sure the date column is an actual date.

Finally, tell me what questions you could ask of this data. Be creative.

**A3** 

```{r}
postpress_url <- "https://www.cardin.senate.gov/?post_type=press-releases"

# extract separately 
pressreleases_url<- read_html("https://www.cardin.senate.gov/?post_type=press-releases")
pressreleases_url %>% html_nodes("date") %>% html_attr('href')
```
```{r}

```

