```{r}
library(tidyverse)
library(janitor)
```


```{r}
# load the  PxP data 
pbp_2023 <- read_csv ("~/Downloads/pbp_2023_d1.csv")
```
```{r}
pbp_2023 %>% 
  group_by(team,action) %>% 
  summarize(total = n())
```
```{r}
pbp_2023 %>% 
  filter(action=="REBOUND", period== 4) 
```

```{r}

```




```{r}
# q1 which big ten team had the most steals in the first quarter?
wbb_pbp <- pbp_2023 %>% 
  #group_by(team,action,type,period) %>% 
  #summarize(total = n()) %>%
  filter(str_detect(conference,"Big Ten")) %>% 
  filter(str_detect(type, "STEAL")) %>% 
  filter(str_detect(period, "1"))  %>% 
group_by(team) %>% 
summarize(total_steals=n()) %>% 
  arrange(desc(total_steals))
```
**A1** Ohio State led the Big Ten in steals with 406 total steals and 11.3 per game, but their data was not featured in the data frame provided. However, Maryland had 192 steals in the first quarter, which led the conference.

```{r}

```





```{r}
# question #2 
# Which team forced the most steals in the fourth quarter all season?
 #group_by(period,team,conference) %>%
  #summarize()
fourq_steals <- pbp_2023 %>%
  filter(str_detect(type, "STEAL")) %>%
  filter(str_detect(period,"4")) %>%
group_by(team)%>%
summarize(total_steals=n()) %>%
  arrange(desc(total_steals))
```
**A2** The Terps collected 158 steals in the fourth quarter.

```{r}
# question #3
# Which team grabbed the most rebounds in the fourth quarter?
rebounds_wbb <- pbp_2023  %>%
filter(str_detect(action,"REBOUND")) %>%
filter(str_detect(period, "4")) %>%
group_by(team)%>%
summarize(total_rebounds=n())%>%
  arrange(desc(total_rebounds))
```
**A3** Maryland grabbed the most rebounds in the fourth quarter with 612.

```{r}
# question 4 
# which Big Ten team made the most three pointers in the third quarter
three_pntrs <- pbp_2023 %>%
filter(str_detect(type, "3PTR")) %>%
filter(str_detect(period, "3")) %>%
filter(str_detect(conference))%>%
group_by(conference)%>%
summarize(total_threes=n())%>%
  arrange(desc(total_threes))
```
**A4** Ironically, Maryland made the most three-pointers in the third quarter with 292.

```{r}
# question 5
# who scored more layups in the fourth quarter: Oklahoma guard Madi Williams or Texas Tech guard Bre'Amber Scott?
# Scott averaged 17.1 ppg, while Williams recorded 15.7 per game. 
oktt_wbb <- pbp_2023 %>%
filter(str_detect(team,"Oklahoma")) %>%
filter(str_detect(player, "25")) %>%
filter(str_detect(type, "LAYUP"))%>%
filter(str_detect(period,"4"))%>%
filter(str_detect(team,"Texas Tech"))%>%
filter(str_detect(player,"23"))%>%
filter(str_detect(type,"LAYUP"))%>%
filter(str_detect(period,"4"))%>%
group_by(player)%>%
summarize(total_layups=n())%>%
  arrange(desc(total_layups))
```
**A5** 

```{r}
# question 6
# which team forced the most turnovers in the second quarter this season
turnovers_wbb <- pbp_2023 %>%
filter(str_detect(action,"TURNOVER")) %>%
filter(str_detect(period,"2"))
group_by(opponent) %>%
summarize(total_turnovers=n())%>%
arrange(desc(total_turnovers))
```

```{r}
# question 7
# how many points three pointers did Brinae Alexander score in the third quarter all season?
umd_alexander <- pbp_2023%>%
filter(str_detect(team,"Maryland"))%>%
filter(str_detect(player, "05")) %>%
filter(str_detect(type,"3PTR")) %>%
filter(str_detect(action,"GOOD")) %>%
filter(str_detect(period,"3")) %>%
group_by(player)%>%
summarize(alexander_threes=n())%>%
  arrange(desc(alexander_threes))
```
**A7** Maryland guard Brinae Alexander made 34 three-pointers in the third quarter.


```{r}
# question 8
# which team committed the most fouls in the first quarter this season?
fouls_total <- pbp_2023 %>%
filter(str_detect(action,"FOUL")) %>%
filter(str_detect(period,"1")) %>%
group_by(team)%>%
summarize(total_fouls=n())%>%
  arrange(desc(total_fouls))
```
**A9** This season,Maryland had the most first quarter fouls with 196.

```

