```{r}
library(tidyverse)
library(janitor)
```


```{r}
# load the  PxP data 
pbp_2023 <- read_csv ("~/Downloads/pbp_2023_d1.csv")
```
```{r}
pbp_2023 %>% 
  group_by(team,action) %>% 
  summarize(total = n())
```
```{r}
pbp_2023 %>% 
  filter(action=="REBOUND", period== 4) 
```

```{r}

```




```{r}
# q1 which big ten team had the most steals in the first quarter?
wbb_pbp <- pbp_2023 %>% 
  #group_by(team,action,type,period) %>% 
  #summarize(total = n()) %>%
  filter(str_detect(conference,"Big Ten")) %>% 
  filter(str_detect(type, "STEAL")) %>% 
  filter(str_detect(period, "1"))  %>% 
group_by(team) %>% 
summarize(total_steals=n()) %>% 
  arrange(desc(total_steals))
```
**A1** Maryland had 192 steals in the first quarter, which led the Big Ten.




```{r}
# question #2 
# Which team forced the most steals in the fourth quarter all season?
fourq_steals <- pbp_2023 %>%
  #group_by(period,team,conference) %>%
  #summarize()
  filter(str_detect(type, "STEAL")) %>%
  filter(str_detect(period,"4")) %>%
group_by(team)%>%
summarize(total_steals=n()) %>%
  arrange(desc(total_steals))
```
**A2** The Terps collected 158 steals in the fourth quarter.

```{r}
# question #3
# Which team grabbed the most rebounds in the fourth quarter?
rebounds_wbb <- pbp_2023  %>%
filter(str_detect(action,"REBOUND")) %>%
filter(str_detect(period, "4")) %>%
group_by(team)%>%
summarize(total_rebounds=n())%>%
  arrange(desc(total_rebounds))
```
**A3** Maryland grabbed the most rebounds in the fourth quarter with 612.

```{r}
# question 4 
# which team made the most three pointers in the third quarter
three_pntrs <- pbp_2023 %>%
filter(str_detect(type, "3PTR")) %>%
filter(str_detect(period, "3")) %>%
group_by(team)%>%
summarize(total_threes=n())%>%
  arrange(desc(total_threes))
```
**A4** Ironically, Maryland made the most three-pointers in the third quarter with 292.

```{r}
# question 5
# how many layups did Baylor guard Aijha Blackwell make in the fourth quarter all season?
baylor_wbb <- pbp_2023 %>%
filter(str_detect(team,"Baylor"))%>%
filter(str_detect(player, "33")) %>%
filter(str_detect(type, "LAYUP"))%>%
group_by(player)%>%
summarize(total_layups=n())%>%
  arrange(desc(total_layups))
```
**A5** Baylor guard Aijha Blackwell made 13 layups in the fourth quarter during the season.

```{r}
# question 6
# which team forced the most turnovers in the second quarter this season
turnovers_wbb <- pbp_2023 %>%
filter(str_detect(action,"TURNOVER")) %>%
filter(str_detect(period,"2"))
group_by(opponent) %>%
summarize(total_turnovers=n())%>%
arrange(desc(total_turnovers))
```

```{r}
# question 7
# how many points three pointers did Brinae Alexander score in the third quarter all season?
umd_alexander <- pbp_2023%>%
filter(str_detect(team,"Maryland"))%>%
filter(str_detect(player, "05")) %>%
filter(str_detect(type,"3PTR")) %>%
filter(str_detect(action,"GOOD")) %>%
filter(str_detect(period,"3")) %>%
group_by(player)%>%
summarize(alexander_threes=n())%>%
  arrange(desc(alexander_threes))
```
**A7** Maryland guard Brinae Alexander made 34 three-pointers in the third quarter.


```{r}
# question 8
# which team committed the most fouls in the first quarter this season?
fouls_total <- pbp_2023 %>%
filter(str_detect(action,"FOUL")) %>%
filter(str_detect(period,"1")) %>%
group_by(team)%>%
summarize(total_fouls=n())%>%
  arrange(desc(total_fouls))
```
**A9** This season,Maryland had the most first quarter fouls with 196.

```

