---
title: "lab_07"
author: "derek willis"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(lubridate)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key if you haven't already. You won't need to install or overwrite it.

```{r echo=FALSE}
census_api_key("3f95bc5c7a6949265b472d95d7a03633eefb898e")
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2021 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2021 5-year ACS and save them to a dataframe

```{r}
acs_variables <- load_variables(2021, "acs5", cache = TRUE)
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2021 5-year estimates? You will need to find the variable for median age first. 
**A1** Baltimore City, Maryland is the Maryland jurisdiction with the lowest median age of 35.7.

```{r}
md <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2021)

md <- md %>% arrange(estimate) 

md
```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50. 
**A2** Worcester County and Talbot County have a median age of at least 50. 

```{r}
md <- md %>% filter(estimate >= 50) 
```

**Q3** We're interested in knowing more about Montgomery County zip codes where overdose calls are coming from and the demographic characteristics they possess. In particular, we want to calculate a rate of calls per 1,000 population. To do this, you'll need to:

1. Load the Montgomery County 911 overdose calls data.
2. Create a new dataframe with the total number of calls for each zip code.
2. Find the ACS variable pertaining to the total 18 and up population (there are several choices here; you want the simplest one).
3. Get from the ACS data from all Maryland zip codes using that variable.
4. Join the ACS data to the Montgomery County zip code totals dataframe and rename the estimate column to a meaningful name you've given the ACS variable.
5. Add a column to the joined dataframe that calculates a rate of calls per 1,000 population.

Which zip code has the highest rate of calls? Which has the lowest? The zip code '20877' has the highest rate of calls. Also, the zip code '20707' has the lowest rate of calls.

**A3** 

```{r}
maryland911_calldata <- read_csv("data/montgomery_2022_overdoses.csv") %>%
  mutate(zip = as.character(zip))
montgomery_total <- maryland911_calldata %>% group_by(zip) %>% 
  summarise(
    total_calls = n()
  )
 zipcodes_18 <- get_acs(geography = "zcta",
              variables = c(medage = "B09021_001"),
              state = "MD",
              year = 2019)
 montgomery_total911 <- montgomery_total %>% left_join(zipcodes_18, by=c('zip'='GEOID')) %>% 
# how to add a column and rename it
     mutate(calls_per= total_calls/estimate*1000) %>% arrange(desc(calls_per))

```

**Q4** Using [CensusReporter.org](https://censusreporter.org/), examine the zip codes with the highest and lowest rates and compare and contrast their demographic profiles. Is there a story here?

**A4**Yes, there’s a story here. The 20707 refers to Laurel, Maryland and that’s been a known area for 911 calls to come from. But, they have the lowest calls_per number out of any zip codes. While, 20877 is the zip code that belongs to Gaithersburg in Montgomery County and is usually low in crime and 911 calls per year. In this situation, Gaithersburg surprisingly had the highest calls_per number.

**Q5** Go back to your joined dataframe from Q3 and add two more columns representing the minimum and maximum populations based on the estimate and margin of error, then calculate per capita rates for each. Do you get the same two zip codes for the highest and lowest as you did before? Does that give you more or less confidence in your answers to Q3?

**A5** 

```{r}
maryland911_calldata <- read_csv("data/montgomery_2022_overdoses.csv") %>%
  mutate(zip = as.character(zip))
montgomery_total <- maryland911_calldata %>% group_by(zip) %>% 
  summarise(
    total_calls = n()
  )
 zipcodes_18 <- get_acs(geography = "zcta",
              variables = c(medage = "B09021_001"),
              state = "MD",
              year = 2019)
 montgomery_total911 <- montgomery_total %>% left_join(zipcodes_18, by=c('zip'='GEOID')) %>% 
# how to add a column and rename it -- Max
     mutate(calls_per= total_calls + moe_product) %>% arrange(desc(calls_per)) %>% 
   montgomerytotal_max <- montgomery_total911 %>% left_join(montgomery_total911, by=c('zip'= 'Max'))%>% 
   
   # how to add a column and rename it -- Min
 mutate(calls_per= total_calls - moe_product) %>% arrange(desc(calls_per)) %>% 
   montgomerytotal_min <- montgomery_total911 %>% left_join(montgomery_total911, by=c('zip'= 'Min'))%>% 
   
     mutate(calls_per= total_calls/estimate*1000) %>% arrange(desc(calls_per))
   
   



```

